### 2022-11-02 draft/draft.Rmd
still writing

---
output:
  # pdf_document:
  bookdown::pdf_document2:
    includes:
      before_body: title.sty
    keep_tex: true
    citation_package: natbib
    number_sections: yes
# output:
#   bookdown::pdf_document2:
#     citation_package: natbib
#     number_citations: yes
# output: html_document
# output: rticles::rss_article
fontsize: 11pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
language:
  label:
    rem: 'Remark '
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{year,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \DeclareMathOperator*{\argmax}{arg\,max}
- \DeclareMathOperator*{\argmin}{arg\,min}
bibliography: bibliography.bib
biblio-style: apalike
abstract: |
  asdf
---

@@

Yet another approach to adapting $K$-means clustering to graph data is to embed the graph to Euclidean space and then apply Lloyd's algorithm on the embedding. 
In this context, a natural choice of embedding is one in which pairwise Euclidean distances between embedding vectors are (approximately) equal to pairwise graph distances between corresponding vertices. 
For example, the combinatorial Laplacian eigenmap is an embedding in which Euclidean distances between pairs of vectors in the embedding are equal to expected commute times between corresponding pairs of vertices in the graph \citep{vonLuxburg2007}. 
\end{example}

@@

In this work, we explore three types of block models, the stochastic block model, the degree corrected block model, and the popularity adjusted block model, as well as a family of latent space models called generalized random dot product graphs. 
The contributions of this work are as follows. 
First, we show that, similar to how the stochastic block model and degree corrected block model are generalized random dot product graphs with specific latent structures, the popularity adjusted block model is also a generalized random dot product graph with a specific structure. 
Using estimators with well-established consistency properties for the generalized random dot product graph, we develop consistent community detection and parameter estimation algorithms for the popularity adjusted block model.

@@

\begin{example}
Recall the UK faculty network in \ref{ex:uk-faculty}. 
If we wish to analyze this as an SBM or DCBM under the GRDPG framework, we can take the ASE of the adjacency matrix and see whether the embedding vectors lie around point masses or rays. 
We will assume that this model is assortative, so $p = K$ and $q = 0$ in constructing the ASE. 
Because the fourth community only contains two members, we omit it from this analysis. 
Figure \ref{fig:ukfaculty-ase} suggests that the appropriate model here is the DCBM. 
Fitting three line segments to these data and clustering the embedding vectors by which line segment is closest results in a community detection accuracy of 83.5\%.

```{r ukfaculty-ase, fig.cap = 'The ASE of the UK faculty network. The vertices are labeled by the school to which each faculty member belongs.', fig.height = 3, fig.width = 4, cache = TRUE }
data(UKfaculty)
A <- as.matrix(igraph::as_adjacency_matrix(UKfaculty, type = 'both'))
A <- sign(A + t(A))
z <- factor(igraph::vertex_attr(UKfaculty)$Group)
A <- A[z != 4, z != 4]
z <- z[z != 4]
Xhat <- embedding(A, 3, 0)
pairs(Xhat, col = z, asp = 1)
```
\end{example}

The PABM is also a Bernoulli random graph, so it is also a GRDPG. 
We explore this in the next section (see section \@ref(sec:pabm-grdpg)).

@@

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for these graphs (figure \ref{fig:mbconnectome-ase}). 
Fitting two quadratic Bezier curves to the ASEs (figure \ref{fig:mbconnectome-kcurves}) results in 73.3\% accuracy for the left hemisphere connectome and 76.5\% accuracy for the right hemisphere connectome.

@@

```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type.', fig.height = 2, fig.width = 8}
Xhat.l <- embedding(A.l, 2, 0)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 0)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

@@

```{r mbconnectome-kcurves, fig.cap = 'ASEs of the Drosophila connectome graphs with quadratic Bezier curves fitted via $K$-curves clustering. The embedding vectors are colored by predicted label.', fig.height = 2, fig.width = 8, cache = TRUE}
set.seed(1)

z.init.r <- sample.points(z.r, 10)
z.init.l <- sample.points(z.l, 10)

>>>>

---
output:
  # pdf_document:
  bookdown::pdf_document2:
    includes:
      before_body: title.sty
    keep_tex: true
    citation_package: natbib
    number_sections: yes
# output:
#   bookdown::pdf_document2:
#     citation_package: natbib
#     number_citations: yes
# output: html_document
# output: rticles::rss_article
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"
urlcolor: blue
language:
  label:
    rem: 'Remark '
header-includes:
- \usepackage{setspace}
- \usepackage{float}
- \usepackage{mathtools}
- \usepackage{natbib}
- \usepackage[linesnumbered,ruled,vlined]{algorithm2e} 
- \setcitestyle{year,comma}
- \usepackage{verbatim}
- \usepackage{amsthm}
- \usepackage{comment}
- \DeclareMathOperator*{\argmax}{arg\,max}
- \DeclareMathOperator*{\argmin}{arg\,min}
bibliography: bibliography.bib
biblio-style: apalike
abstract: |
  asdf
---

@@

Yet another approach to adapting $K$-means clustering to graph data is to embed the graph to Euclidean space and then apply Lloyd's algorithm on the embedding. 
In this context, a natural choice of embedding is one in which pairwise Euclidean distances between embedding vectors are (approximately) equal to pairwise graph distances between corresponding vertices. 
For example, the combinatorial Laplacian eigenmap is an embedding in which Euclidean distances between pairs of vectors in the embedding are equal to expected commute times between corresponding pairs of vertices in the graph \citep{vonLuxburg2007}. 
This embedding is constructed by computing the combinatorial Laplacian matrix $L = D - A$ in which $D$ is a diagonal matrix in which each $D_{ii}$ is the degree of vertex $i$. 
Then the components of the embedding consist of $\big[ v_{n-1} / \sqrt{\lambda_{n - 1}} \mid \cdots \mid v_1 / \sqrt{\lambda_1} \big]$ in which $\lambda_i$ is the $i^{th}$ largest eigenvalue of $L$ and $v_i$ is its corresponding eigenvector ($\lambda_n = 0$ and is omitted). 
Then Lloyd's algorithm can be applied to this embedding. 
More often in practice, the $d$ smallest eigenvalues are used to construct a lower dimensional embedding. 
\end{example}

@@

In this work, we explore three types of block models, the stochastic block model, the degree corrected block model, and the popularity adjusted block model, as well as a family of latent space models called generalized random dot product graphs.

The contributions of this work are as follows.

In section \@ref(sec:pabm-grdpg), we show that, similar to how the stochastic block model and degree corrected block model are generalized random dot product graphs with specific latent structures, the popularity adjusted block model is also a generalized random dot product graph with a specific structure. 
More specifically, we show that the popularity adjusted block model is a generalized random dot product graph in which the latent vectors are organized by community and each community in the latent space is defined by a $K$-dimensional subspace, where $K$ is the number of communities in the model. 
We then exploit this rigid and linear structure for community detection by applying sparse subspace clustering on the adjacency spectral embedding of the graph to infer the community labels. 
We also propose a novel algorithm called orthogonal spectral clustering, which constructs an embedding for the popularity adjusted block model which forces the communities to lie on *orthogonal* subspaces. 
Finally, we propose a spectral decomposition based parameter estimation algorithm for the popularity adjusted block model. 
All three algorithms have consistency properties that result in zero error with increasing graph size.

In section \@ref(sec:mbm), we extend the latent structures of the stochastic block model, degree corrected block model, and popularity adjusted block model to nonlinear latent structures. 
Instead of points, line segments, and subspaces, we propose the manifold block model which is a class of generalized random dot product graph in which the latent vectors lie on one of $K$ manifolds. 
In the case of nonintersecting manifolds, we show that a clustering algorithm based on a nearest algorithms graph approach on an embedding is consistent, resulting in zero clustering error for large graphs, as long as the latent vectors are sampled from the manifolds densely 
In the case of intersecting manifolds, we propose the $K$-curves algorithm, which is a type of batch coordinate ascent algorithm which alternates between fitting manifolds to the embedding and reassigning clusters based on proximity to the manifolds.

@@

The latent structures of the SBM and DCBM when viewed as GRDPGs lend themselves to convenient community detection algorithms based on existing clustering algorithms for Euclidean data. 
The general strategy for these models is to construct the appropriate ASE, choosing the number of components based on the number of communities and whether to use positive or negative eigenvalues depending on assortativity or disassortativity, and then to apply the appropriate Euclidean-based clustering algorithm for the specific structure. 
In the case of the SBM, after embedding in $K$ dimensions, since the original latent structure consists of $K$ point masses, an appropriate clustering algorithm might be one based on centroids, such as $K$-means clustering or Gaussian mixture models. 
In the case of the DCBM, after embedding in $K$ dimensions, since the original latent structure consists of $K$ rays emanating from the origin, an appropriate clustering algorithm is one based on angles between points rather than distances, such as kernel $K$-means clustering using the cosine kernel. 
Another clustering algorithm for the ASE of a DCBM is to normalize the embedding vectors to norm 1 and then apply $K$-means to the normalized vectors.

\begin{example}
Recall the UK faculty network in \ref{ex:uk-faculty}. 
If we wish to analyze this as an SBM or DCBM under the GRDPG framework, we can take the ASE of the adjacency matrix and see whether the embedding vectors lie around point masses or rays. 
We assume that this model is assortative, so $p = K$ and $q = 0$ in constructing the ASE. 
Because the fourth community only contains two members, we omitted it from this analysis. 
Figure \ref{fig:ukfaculty-ase} suggests that the appropriate model here is the DCBM. 
$K$-means clustering on the normalized embedding vectors results in a community detection accuracy of 96.7\%.

```{r ukfaculty-ase, fig.cap = 'A three-dimensional ASE of the UK faculty network using the three most positive eigenvalues and their corresponding eigenvectors. The vertices are labeled by the school to which each faculty member belongs.', fig.height = 3, fig.width = 5, cache = TRUE }
data(UKfaculty, package = 'igraphdata')
A <- as.matrix(igraph::as_adjacency_matrix(UKfaculty, type = 'both'))
A <- sign(A + t(A))
z <- factor(igraph::vertex_attr(UKfaculty)$Group)
A <- A[z != 4, z != 4]
z <- z[z != 4]
Xhat <- embedding(A, 3, 0)
# pairs(Xhat, col = z, asp = 1)
Xhat %>% 
  as.data.frame %>% 
  magrittr::set_colnames(c('comp 1', 'comp 2', 'comp 3')) %>% 
  GGally::ggpairs(aes(colour = z)) + 
  scale_colour_brewer(palette = 'Set1') + 
  theme_bw()
```

The PABM is also a Bernoulli random graph, so it is also a GRDPG. 
The latent vectors of the PABM when viewed as a GRDPG form $K$ $K$-dimensional subspaces, which makes for an elegant parallel between the nested forms of the block models and the nested structures of the GRDPGs: The SBM consists of $K$ zero-dimensional subspaces, and DCBM consists of $K$ one-dimensional subspaces. 
We explore this connection further the next section (section \@ref(sec:pabm-grdpg)).

@@

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for these graphs (figure \ref{fig:mbconnectome-ase}). 
Fitting two quadratic Bezier curves to the ASEs (figure \ref{fig:mbconnectome-kcurves}) results in 72.2\% accuracy for the left hemisphere connectome and 83.6\% accuracy for the right hemisphere connectome. 
The curves were fitted on three-dimensional ASEs with $p = 2$ and $q = 1$, which were chosen based on setting the threshold $|\lambda_i| > \sqrt{n}$ where $\lambda_i$ is the $i$^th^ greatest eigenvalue. 
Figure \ref{fig:mbconnectome-kcurves} shows the embedding and curves projected to the first two components.

@@

```{r mbconnectome-ase, fig.cap = 'ASEs of the Drosophila connectome graphs. The embedding vectors are labeled by neuron type.', fig.height = 2, fig.width = 8}
Xhat.l <- embedding(A.l, 2, 1)
ase.l <- ggplot() + 
  geom_point(aes(x = Xhat.l[, 1], y = Xhat.l[, 2],
                 colour = factor(z.l))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

Xhat.r <- embedding(A.r, 2, 1)
ase.r <- ggplot() + 
  geom_point(aes(x = Xhat.r[, 1], y = Xhat.r[, 2],
                 colour = factor(z.r))) + 
  coord_fixed(.5) +
  theme_bw() + 
  theme(legend.position = 'none') +
  scale_colour_brewer(palette = 'Set1') + 
  labs(x = NULL, y = NULL, colour = NULL)

@@

```{r mbconnectome-kcurves, fig.cap = 'ASEs of the Drosophila connectome graphs with quadratic Bezier curves fitted via $K$-curves clustering. The embedding vectors are colored by predicted label.', fig.height = 2, fig.width = 8, cache = TRUE}
set.seed(123456)

z.init.r <- sample.points(z.r, 6)
z.init.l <- sample.points(z.l, 6)

### 2022-11-02 draft/draft.tex
still writing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{year,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{apalike}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

@@

In this work, we explore three types of block models, the stochastic block model, the degree corrected block model, and the popularity adjusted block model, as well as a family of latent space models called generalized random dot product graphs.
The contributions of this work are as follows.
First, we show that, similar to how the stochastic block model and degree corrected block model are generalized random dot product graphs with specific latent structures, the popularity adjusted block model is also a generalized random dot product graph with a specific structure.
Using estimators with well-established consistency properties for the generalized random dot product graph, we develop consistent community detection and parameter estimation algorithms for the popularity adjusted block model.

@@

\begin{example}
Recall the UK faculty network in \ref{ex:uk-faculty}. 
If we wish to analyze this as an SBM or DCBM under the GRDPG framework, we can take the ASE of the adjacency matrix and see whether the embedding vectors lie around point masses or rays. 
We will assume that this model is assortative, so $p = K$ and $q = 0$ in constructing the ASE. 
Because the fourth community only contains two members, we omit it from this analysis. 
Figure \ref{fig:ukfaculty-ase} suggests that the appropriate model here is the DCBM. 
Fitting three line segments to these data and clustering the embedding vectors by which line segment is closest results in a community detection accuracy of 83.5\%.

@@

\caption{The ASE of the UK faculty network. The vertices are labeled by the school to which each faculty member belongs.}\label{fig:ukfaculty-ase}
\end{figure}
\end{example}

The PABM is also a Bernoulli random graph, so it is also a GRDPG.
We explore this in the next section (see section \ref{sec:pabm-grdpg}).

@@

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for these graphs (figure \ref{fig:mbconnectome-ase}).
Fitting two quadratic Bezier curves to the ASEs (figure \ref{fig:mbconnectome-kcurves}) results in 78.5\% accuracy for the left hemisphere connectome and 74.6\% accuracy for the right hemisphere connectome.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{year,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{apalike}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

@@

In this work, we explore three types of block models, the stochastic block model, the degree corrected block model, and the popularity adjusted block model, as well as a family of latent space models called generalized random dot product graphs.

The contributions of this work are as follows.

In section \ref{sec:pabm-grdpg}, we show that, similar to how the stochastic block model and degree corrected block model are generalized random dot product graphs with specific latent structures, the popularity adjusted block model is also a generalized random dot product graph with a specific structure.
More specifically, we show that the popularity adjusted block model is a generalized random dot product graph in which the latent vectors are organized by community and each community in the latent space is defined by a \(K\)-dimensional subspace, where \(K\) is the number of communities in the model.
We then exploit this rigid and linear structure for community detection by applying sparse subspace clustering on the adjacency spectral embedding of the graph to infer the community labels.
We also propose a novel algorithm called orthogonal spectral clustering, which constructs an embedding for the popularity adjusted block model which forces the communities to lie on \emph{orthogonal} subspaces.
Finally, we propose a spectral decomposition based parameter estimation algorithm for the popularity adjusted block model.
All three algorithms have consistency properties that result in zero error with increasing graph size.

In section \ref{sec:mbm}, we extend the latent structures of the stochastic block model, degree corrected block model, and popularity adjusted block model to nonlinear latent structures.
Instead of points, line segments, and subspaces, we propose the manifold block model which is a class of generalized random dot product graph in which the latent vectors lie on one of \(K\) manifolds.
In the case of nonintersecting manifolds, we show that a clustering algorithm based on a nearest algorithms graph approach on an embedding is consistent, resulting in zero clustering error for large graphs, as long as the latent vectors are sampled from the manifolds densely
In the case of intersecting manifolds, we propose the \(K\)-curves algorithm, which is a type of batch coordinate ascent algorithm which alternates between fitting manifolds to the embedding and reassigning clusters based on proximity to the manifolds.

@@

The latent structures of the SBM and DCBM when viewed as GRDPGs lend themselves to convenient community detection algorithms based on existing clustering algorithms for Euclidean data.
The general strategy for these models is to construct the appropriate ASE, choosing the number of components based on the number of communities and whether to use positive or negative eigenvalues depending on assortativity or disassortativity, and then to apply the appropriate Euclidean-based clustering algorithm for the specific structure.
In the case of the SBM, after embedding in \(K\) dimensions, since the original latent structure consists of \(K\) point masses, an appropriate clustering algorithm might be one based on centroids, such as \(K\)-means clustering or Gaussian mixture models.
In the case of the DCBM, after embedding in \(K\) dimensions, since the original latent structure consists of \(K\) rays emanating from the origin, an appropriate clustering algorithm is one based on angles between points rather than distances, such as kernel \(K\)-means clustering using the cosine kernel.
Another clustering algorithm for the ASE of a DCBM is to normalize the embedding vectors to norm 1 and then apply \(K\)-means to the normalized vectors.

\begin{example}
Recall the UK faculty network in \ref{ex:uk-faculty}. 
If we wish to analyze this as an SBM or DCBM under the GRDPG framework, we can take the ASE of the adjacency matrix and see whether the embedding vectors lie around point masses or rays. 
We assume that this model is assortative, so $p = K$ and $q = 0$ in constructing the ASE. 
Because the fourth community only contains two members, we omitted it from this analysis. 
Figure \ref{fig:ukfaculty-ase} suggests that the appropriate model here is the DCBM. 
$K$-means clustering on the normalized embedding vectors results in a community detection accuracy of 96.7\%.

@@

\caption{A three-dimensional ASE of the UK faculty network using the three most positive eigenvalues and their corresponding eigenvectors. The vertices are labeled by the school to which each faculty member belongs.}\label{fig:ukfaculty-ase}
\end{figure}

The PABM is also a Bernoulli random graph, so it is also a GRDPG.
The latent vectors of the PABM when viewed as a GRDPG form \(K\) \(K\)-dimensional subspaces, which makes for an elegant parallel between the nested forms of the block models and the nested structures of the GRDPGs: The SBM consists of \(K\) zero-dimensional subspaces, and DCBM consists of \(K\) one-dimensional subspaces.
We explore this connection further the next section (section \ref{sec:pabm-grdpg}).

@@

The ASE in two dimensions suggest that the MBM with one dimensional quadratic manifolds is an appropriate model for these graphs (figure \ref{fig:mbconnectome-ase}).
Fitting two quadratic Bezier curves to the ASEs (figure \ref{fig:mbconnectome-kcurves}) results in 72.2\% accuracy for the left hemisphere connectome and 83.6\% accuracy for the right hemisphere connectome.
The curves were fitted on three-dimensional ASEs with \(p = 2\) and \(q = 1\), which were chosen based on setting the threshold \(|\lambda_i| > \sqrt{n}\) where \(\lambda_i\) is the \(i\)\textsuperscript{th} greatest eigenvalue.
Figure \ref{fig:mbconnectome-kcurves} shows the embedding and curves projected to the first two components.
